---
title: "NewGlobe Case Study - Analyst M&E/Data Analytics Teams"
author: "Kamran Ahmed"
date: "2023-10-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The Data
You have received four files, all in .dta and .xlsx formats, so you can use whichever format you prefer. These files are the following:

● “Lesson completion”: file provided at the teacher level, meaning that there is a unique row for each teacher. The file contains the grade that each teacher teaches, and the average lesson completion rate over the term of interest.

● “Pupil attendance”: file provided at the pupil level (that means that there is a unique row for each pupil). This file includes the unique school ID, unique pupil ID, the pupil’s grade, the attendance records, and the present records.

○ The attendance records means the total number of times that a pupil’s teacher took attendance.

○ The present records means the total number of times that a pupil was present, out of the attendance


● “Pupil scores”: file provided at the pupil*subject level (that means that there are more than one row per pupil). This file includes the unique school ID, unique pupil ID, the pupil’s grade, the subject for this assessment, and the score obtained in this assessment.

● “School information”: file provided at the school-level. It includes the region and province where each school is located, the unique school ID, and the “treatment status” (yes/no) for a given tutoring program.

# Load libraries
```{r}
library(tidyverse)
library(readxl)
```
# Load Data

Let's load the data and have a quick look of it to see how the data look like
```{r}
Lesson_completion <- read_excel("Lesson completion.xlsx")
head(Lesson_completion)
```

```{r}
Pupil_attendance <- read_excel("Pupil attendance.xlsx")
head(Pupil_attendance)
```

```{r}
Pupil_scores <- read_excel("Pupil scores.xlsx")
head(Pupil_scores)
```

```{r}
School_information <- read_excel("School_information.xlsx")
head(School_information)
```

# Step 1: Data cleaning

Please create a file at the student-level which has information about their test scores, school information, their attendance, and their teacher’s lesson completion rate. Note that this is the main data set that we expect you to share with us.

Hint: note that the four data sets you will use are all presented at different “levels” of the data (e.g., “School information” is at the level of the school, but “Pupil scores” is at the level of the student). Therefore, we suggest that you start by reshaping the “Pupil scores” file so that each student only has one row in the data, with different columns for their scores in math, fluency, and Kiswahili. Use this as your “base file”, and start merging all the other files to this. Be careful with how you merge things: since there are many students to a school or even a teacher, some of these merges will need to be “many-to-one” (but not all).
```{r}
# Create a base file by reshaping the “Pupil scores” file so that each student only has one row in the data, with different columns for their scores in math, fluency, and Kiswahili
Pupil_scores <- Pupil_scores %>%
  pivot_wider(names_from = subject, values_from = score)
head(Pupil_scores)
```

```{r}
# Merge Pupil_attendance file to the base file i.e, Pupil_scores file
pupil_df <- merge(Pupil_scores, Pupil_attendance, by = c("pupil_id", "school_id", "grade"))
head(pupil_df)
```

```{r}
# Merge teacher's data
pupil_teacher_df <- merge(pupil_df, Lesson_completion, by = c("school_id", "grade"))
head(pupil_teacher_df)
```

```{r}
#merge school information
pupil_teacher_school_df <- merge(pupil_teacher_df, School_information, by = "school_id")
head(pupil_teacher_school_df)
```

This is the the main data set that we will work with. Let's export this as a csv file and name it "main_data"
```{r}
write.csv(pupil_teacher_school_df, file = "main_data.csv", row.names = FALSE)
```


# Step 2: Calculating KPIs

One of our main KPIs within the Schools Vertical is “Percent Pupils Present”. The “layman’s definition” of this KPI is “The percentage of pupils who were present, out of all pupils - across all days in the term to date ”. In other words, the percentage of pupils who were present (for each pupil in the “Pupil attendance” file, this is displayed in the “present_records” variable), out of pupils who had attendance records (the “attendance_records” variable in the same file).

● The first task is to translate this KPI into the data. We will calculate this KPI in two different ways. First, calculate this KPI for all pupils at once. What is the network-level average Percent Pupils Present (use two decimal points)?
```{r}
# Network-Level Average Percent Pupils Present (All Pupils)
network_level_average_kpi <- round(sum(pupil_teacher_school_df$present_records)/sum(pupil_teacher_school_df$attendance_records), 2)
network_level_average_kpi
```

● Now, please calculate this percentage for each school, and create an average at the school-level. What is the average Percent Pupils Present now (use two decimal points)?
```{r}
# School-Level Average Percent Pupils Present
school_level_average_kpi <- pupil_teacher_school_df %>%
  select(school_id, present_records, attendance_records) %>%
  group_by(school_id) %>%
  summarise(total_present = sum(present_records), total_records = sum(attendance_records)) %>%
  mutate(school_kpi = round(total_present/total_records, 2)) %>%
  summarise(round(mean(school_kpi), 2))
school_level_average_kpi
```

● How does the interpretation of the KPI change between the two approaches? Does it matter in this case? When would it matter, (i.e., when would one be more appropriate than the other?) 2-4 sentences max.
```{r}
# The way we interpret the KPI shifts with these two approaches due to their scope. When we calculate the network-level average, we're looking at the big picture, assessing how well the entire network (Bridge Kenya programme) is performing by considering all pupils across all schools. On the other hand, the school-level average narrows our focus to individual school performance, helping us pinpoint differences between schools. In this case it doesn't matter much because we get the same values for the KPIs through both approaches when rounded to two decimal places. However, it would matter when there is more heterogeneity across schools in terms of attendance rate and number of pupils. If there is more variation in attendance rate across schools and number of pupils, merely calculating percentage for each school and creating a simple average at the school-level would give the same weightage to each scool regardless of the number of pupils in that school, hence the value will deviate from the network-level average. The choice between these approaches hinges on the specific analysis or decision-making context. The network-level approach would one be more appropriate when we want to gauge the overall network performance, whereas the school-level approach is valuable for recognizing variations and addressing specific issues within each school. Ultimately, the choice depends on the specific objectives of the analysis or decision-making process.
```

# Step 3: Descriptives

Let’s dig into the reading fluency scores in your current data set. These came from the “Pupil scores” data, but you will need the data set you created in Step 1 above to answer these questions. Please answer the following questions as succinctly as possible.

Please create a figure or a table, whichever you prefer, which shows average fluency scores for each of the five grades.
```{r}
pupil_teacher_school_df %>%
  select(grade, Fluency) %>%
  group_by(grade) %>%
  summarise(mean(Fluency, na.rm = TRUE))
```
```{r}
pupil_teacher_school_df %>%
  select(grade, Fluency) %>%
  group_by(grade) %>%
  summarise(avg_fluency = mean(Fluency, na.rm = TRUE)) %>%
  ggplot(aes(x = grade, y=avg_fluency))+
  geom_col(fill= "dark blue")+
  labs(title = "Average Fluency Scores Across Grades",
       x = "Grade",
       y = "Average Fluency",
       caption = "Based on data data from Bridge Kenya programme")+
  theme(plot.title = element_text(hjust = 0.5))
```

● Which regions (using the “region” variable) have the lowest and highest average fluency score across all grades?
```{r}
pupil_teacher_school_df %>%
  select(region, Fluency) %>%
  group_by(region) %>%
  summarise(avg_fluency = mean(Fluency, na.rm = TRUE)) %>%
  filter(avg_fluency == max(avg_fluency, na.rm=TRUE) | avg_fluency == min(avg_fluency, na.rm=TRUE))
```
Kirinyaga has the lowest average fluency score across all grades.
Machakos has the highest average fluency score across all grades.

● Please create a binary variable that is 1 if a given child reads at 10 or lower, and 0 otherwise. Please create a bar chart with grades on the x-axis, and the share of pupils scoring under this threshold for each grade.
```{r}
pupil_teacher_school_df$not_fluent <- ifelse(pupil_teacher_school_df$Fluency <=10, 1, 0)

pupil_teacher_school_df %>%
  select(grade, not_fluent) %>%
  group_by(grade) %>%
  summarise(proportion_not_fluent = mean(not_fluent, na.rm = TRUE)) %>%
  ggplot(aes(x = grade, y=proportion_not_fluent*100))+
  geom_col(fill= "dark blue")+
  labs(title = "Share of pupils scoring under Reading Fluency threshold of 10 across Grades",
       x = "Grade",
       y = "% of Pupils below threshold",
       caption = "Based on data data from Bridge Kenya programme")+
  theme(plot.title = element_text(hjust = 0.5))
```

● What school has the highest share of pupils scoring under this threshold in grade 3?
```{r}
pupil_teacher_school_df %>%
  filter(grade == "Grade 3") %>%
  select(school_id, not_fluent) %>%
  group_by(school_id) %>%
  summarise(proportion_not_fluent = mean(not_fluent, na.rm = TRUE)) %>%
  filter(proportion_not_fluent == max(proportion_not_fluent, na.rm=TRUE))
```
school_id 223941 is the one that has the highest share of pupils scoring under this threshold in grade 3.

# Step 4: Impact evaluation 

During this term, we rolled out an intensive after-school tutoring program in 55 schools. The selection to be a part of the 55 schools was randomly assigned - in other words, these schools were part of a randomized controlled trial (RCT). The “School Information” data set has a binary variable for whether each school was part of the program or not.

● Our Chief Academic Officer would like to know whether this program had any effects on test scores in math, Kiswahili, fluency, and/or student attendance. Please conduct any calculations you see fit to answer his questions.

Let's do the following calculations to transform the columns first and prepare the data for regression analysis.
```{r}
# Create a column for attendance performance at the student level
pupil_teacher_school_df$student_attendance_score <- 100*(pupil_teacher_school_df$present_records/pupil_teacher_school_df$attendance_records)

# The column for Math, Kiswahili and lesson completion rate currently have values between 0 and 1. Let's multiple them by 100 to show them in percentage form 

pupil_teacher_school_df$Math <- pupil_teacher_school_df$Math*100
pupil_teacher_school_df$Kiswahili <- pupil_teacher_school_df$Kiswahili*100
pupil_teacher_school_df$lesson_completion_rate <- pupil_teacher_school_df$lesson_completion_rate*100
```

```{r}
summary(lm(Math ~ tutoring_program, data = pupil_teacher_school_df))
```
```{r}
summary(lm(Kiswahili ~ tutoring_program, data = pupil_teacher_school_df))
```
```{r}
summary(lm(Fluency ~ tutoring_program, data = pupil_teacher_school_df))
```

```{r}
summary(lm(student_attendance_score ~ tutoring_program, data = pupil_teacher_school_df))
```
```{r}
# By running simple linear regression with Math, Kiswahili, Fluency, and Attendance score as the dependent variable and tutoring program as the independent variable we saw that tutoring program has a statistically significant positive effect on students' test scores in Math, Kiswahili, Fluency, and attendance on average. The results are highly statistically significant even at a significance level as low as 0.001.
```


● After conducting the impact evaluation, we have heard anecdotally that teachers in schools that received tutoring felt more motivated and were completing their lessons at a faster pace. Hence, we could worry that the effects that we see are not (solely) due to the tutoring program, but also due to the higher lesson completion rate. Does this hypothesis hold up in the data?

```{r}
# If teachers in schools that received tutoring indeed felt more motivated and were completing their lessons at a faster pace then tutoring program also has an effect on teachers completion rate. In this case teacher's lesson completion rate is a confounding variable that also effects the outcome variable and is being effected by the treatment variable (tutoring program). If this is true, omitting teachers lesson completion rate from the regressions would make the results bias as the effects of teachers higher lesson completion rate would also be wrongly attributed to the tutoring program. Let's first run a regression of teachers lesson completion rate on tutoring program to see if there is any effect of tutoring program on lesson competion rate. Then, we will modify the above regressions by controlling for the effects of lesson completion rate by introducing the lesson_completion_rate to the regression along with the tutoring program.
```

```{r}
summary(lm(lesson_completion_rate ~ tutoring_program, data = pupil_teacher_school_df))
```
```{r}
# In the above regression results, we see that tutoring program does not have a statistically significant effect on lesson completion rate. Therefore, we do not need to worry that the effects that we see are not (solely) due to the tutoring program, but also due to the higher lesson completion rate.
# Below, we control for lesson_completion_rate to see if te results of regression vary from the above ones
```


```{r}
summary(lm(Math ~ tutoring_program+lesson_completion_rate, data = pupil_teacher_school_df))
```
```{r}
summary(lm(Kiswahili ~ tutoring_program+lesson_completion_rate, data = pupil_teacher_school_df))
```

```{r}
summary(lm(Fluency ~ tutoring_program+lesson_completion_rate, data = pupil_teacher_school_df))
```

```{r}
summary(lm(student_attendance_score ~ tutoring_program+lesson_completion_rate, data = pupil_teacher_school_df))
```


