---
title: "NewGlobe Case Study - Analyst M&E/Data Analytics Teams"
author: "Kamran Ahmed"
date: "2023-10-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries
```{r}
library(tidyverse)
library(readxl)
```
# Load Data
```{r}
Lesson_completion <- read_excel("Lesson completion.xlsx")
head(Lesson_completion)
```

```{r}
Pupil_attendance <- read_excel("Pupil attendance.xlsx")
head(Pupil_attendance)
```

```{r}
Pupil_scores <- read_excel("Pupil scores.xlsx")
head(Pupil_scores)
```

```{r}
School_information <- read_excel("School_information.xlsx")
head(School_information)
```

# Step 1: Data cleaning

Please create a file at the student-level which has information about their test scores, school information, their attendance, and their teacher’s lesson completion rate. Note that this is the main data set that we expect you to share with us.

Hint: note that the four data sets you will use are all presented at different “levels” of the data (e.g., “School information” is at the level of the school, but “Pupil scores” is at the level of the student). Therefore, we suggest that you start by reshaping the “Pupil scores” file so that each student only has one row in the data, with different columns for their scores in math, fluency, and Kiswahili. Use this as your “base file”, and start merging all the other files to this. Be careful with how you merge things: since there are many students to a school or even a teacher, some of these merges will need to be “many-to-one” (but not all).
```{r}
# Create a base file by reshaping the “Pupil scores” file
Pupil_scores <- Pupil_scores %>%
  pivot_wider(names_from = subject, values_from = score)
head(Pupil_scores)
```

```{r}
# Merge Pupil_attendance file to the base file i.e, Pupil_scores file
pupil_df <- merge(Pupil_scores, Pupil_attendance, by = c("pupil_id", "school_id", "grade"))
head(pupil_df)
```

```{r}
# Merge teacher's data
pupil_teacher_df <- merge(pupil_df, Lesson_completion, by = c("school_id", "grade"))
head(pupil_teacher_df)
```

```{r}
#merge school information
pupil_teacher_school_df <- merge(pupil_teacher_df, School_information, by = "school_id")
head(pupil_teacher_school_df)
```

# Step 2: Calculating KPIs

One of our main KPIs within the Schools Vertical is “Percent Pupils Present”. The “layman’s definition” of this KPI is “The percentage of pupils who were present, out of all pupils - across all days in the term to date ”. In other words, the percentage of pupils who were present (for each pupil in the “Pupil attendance” file, this is displayed in the “present_records” variable), out of pupils who had attendance records (the “attendance_records” variable in the same file).

● The first task is to translate this KPI into the data. We will calculate this KPI in two different ways. First, calculate this KPI for all pupils at once. What is the network-level average Percent Pupils Present (use two decimal points)?
```{r}
network_level_average_kpi <- round(sum(pupil_teacher_school_df$present_records)/sum(pupil_teacher_school_df$attendance_records), 2)
network_level_average_kpi
```

● Now, please calculate this percentage for each school, and create an average at the school-level. What is the average Percent Pupils Present now (use two decimal points)?
```{r}
school_level_average_kpi <- pupil_teacher_school_df %>%
  select(school_id, present_records, attendance_records) %>%
  group_by(school_id) %>%
  summarise(total_present = sum(present_records), total_records = sum(attendance_records)) %>%
  mutate(school_kpi = round(total_present/total_records, 2)) %>%
  summarise(round(mean(school_kpi), 2))
school_level_average_kpi
```

● How does the interpretation of the KPI change between the two approaches? Does it matter in this case? When would it matter, (i.e., when would one be more appropriate than the other?) 2-4 sentences max.
```{r}

```

# Step 3: Descriptives

Let’s dig into the reading fluency scores in your current data set. These came from the “Pupil scores” data, but you will need the data set you created in Step 1 above to answer these questions. Please answer the following questions as succinctly as possible.

Please create a figure or a table, whichever you prefer, which shows average fluency scores for each of the five grades.
```{r}
pupil_teacher_school_df %>%
  select(grade, Fluency) %>%
  group_by(grade) %>%
  summarise(mean(Fluency, na.rm = TRUE))
```
```{r}
pupil_teacher_school_df %>%
  select(grade, Fluency) %>%
  group_by(grade) %>%
  summarise(avg_fluency = mean(Fluency, na.rm = TRUE)) %>%
  ggplot(aes(x = grade, y=avg_fluency))+
  geom_col(fill= "dark blue")+
  labs(title = "Average Fluency Scores for Grades",
       x = "Grade",
       y = "Average Fluency",
       caption = "Based on data data from Bridge Kenya programme")+
  theme(plot.title = element_text(hjust = 0.5))
```

● Which regions (using the “region” variable) have the lowest and highest average fluency score across all grades?
```{r}
pupil_teacher_school_df %>%
  select(region, Fluency) %>%
  group_by(region) %>%
  summarise(avg_fluency = mean(Fluency, na.rm = TRUE)) %>%
  filter(avg_fluency == max(avg_fluency, na.rm=TRUE) | avg_fluency == min(avg_fluency, na.rm=TRUE))
```

● Please create a binary variable that is 1 if a given child reads at 10 or lower, and 0 otherwise. Please create a bar chart with grades on the x-axis, and the share of pupils scoring under this threshold for each grade.
```{r}
pupil_teacher_school_df$not_fluent <- ifelse(pupil_teacher_school_df$Fluency <=10, 1, 0)

pupil_teacher_school_df %>%
  select(grade, not_fluent) %>%
  group_by(grade) %>%
  summarise(proportion_not_fluent = mean(not_fluent, na.rm = TRUE)) %>%
  ggplot(aes(x = grade, y=proportion_not_fluent))+
  geom_col(fill= "dark blue")+
  labs(title = "Share of pupils scoring under Reading Fluency threshold of 10 across Grades",
       x = "Grade",
       y = "Proportion of Pupils below threshold",
       caption = "Based on data data from Bridge Kenya programme")+
  theme(plot.title = element_text(hjust = 0.5))
```

● What school has the highest share of pupils scoring under this threshold in grade 3?
```{r}
pupil_teacher_school_df %>%
  filter(grade == "Grade 3") %>%
  select(school_id, not_fluent) %>%
  group_by(school_id) %>%
  summarise(proportion_not_fluent = mean(not_fluent, na.rm = TRUE)) %>%
  filter(proportion_not_fluent == max(proportion_not_fluent, na.rm=TRUE))
```

# Step 4: Impact evaluation 

During this term, we rolled out an intensive after-school tutoring program in 55 schools. The selection to be a part of the 55 schools was randomly assigned - in other words, these schools were part of a randomized controlled trial (RCT). The “School Information” data set has a binary variable for whether each school was part of the program or not.

● Our Chief Academic Officer would like to know whether this program had any effects on test scores in math, Kiswahili, fluency, and/or student attendance. Please conduct any calculations you see fit to answer his questions.
```{r}
summary(lm(Math ~ tutoring_program, data = pupil_teacher_school_df))
```
```{r}
summary(lm(Kiswahili ~ tutoring_program, data = pupil_teacher_school_df))
```
```{r}
summary(lm(Fluency ~ tutoring_program, data = pupil_teacher_school_df))
```

```{r}
pupil_teacher_school_df$student_attendance_score <- pupil_teacher_school_df$present_records/pupil_teacher_school_df$attendance_records

summary(lm(student_attendance_score ~ tutoring_program, data = pupil_teacher_school_df))

```

● After conducting the impact evaluation, we have heard anecdotally that teachers in schools that received tutoring felt more motivated and were completing their lessons at a faster pace. Hence, we could worry that the effects that we see are not (solely) due to the tutoring program, but also due to the higher lesson completion rate. Does this hypothesis hold up in the data?
```{r}
summary(lm(Math ~ tutoring_program+lesson_completion_rate, data = pupil_teacher_school_df))
```
```{r}
summary(lm(Kiswahili ~ tutoring_program+lesson_completion_rate, data = pupil_teacher_school_df))
```

```{r}
summary(lm(Fluency ~ tutoring_program+lesson_completion_rate, data = pupil_teacher_school_df))
```

```{r}
summary(lm(student_attendance_score ~ tutoring_program+lesson_completion_rate, data = pupil_teacher_school_df))
```


